Associahedra for finite type cluster algebras \\ and minimal relations between $\b{g}$-vectors

### 2020-05-05 thirdPaper/draft/typeConeAssociahedra.tex
footnote graph associahedra

\smallskip
Further families of generalizations of the associahedron certainly deserve to be studied with our type cone approach.
We are investigating in particular graph associahedra~\cite{CarrDevadoss, Postnikov, FeichtnerSturmfels, Zelevinsky}, brick polytopes~\cite{PilaudSantos-brickPolytope, PilaudStump-brickPolytope} and quotientopes~\cite{Reading-latticeCongruences, Reading-CambrianLattices, PilaudSantos-quotientopes}.
In fact, as already mentioned, it seems sometimes easier to find all polytopal realizations of a fan (by describing its type cone in terms of certain specific pairs of adjacent maximal cones) than to identify one specific explicit realization.
This is clearly the case when the type cone is simplicial as illustrated by the results of this paper: we naturally obtained all polytopal realizations of the $\b{g}$-vector fans of cluster algebras from cyclic seeds and of gentle algebras whose polytopality was only established very recently~\cite{HohlwegPilaudStella, PaluPilaudPlamondon-nonkissing}.
We believe that this approach gives reasonable hope that the polytopality of further fans could in the future be established using our type cone approach.
Tempting candidates are quotientopes for hyperplane arrangements beyond type~$A$, which are particularly interesting since the coarsenings of a fan essentially correspond to the faces of its type cone (see \cref{subsec:facesTypeCone}).

@@

\para{Principal coefficients and $\b{g}$- and $\b{c}$-vectors}
%
We now consider principal coefficients to define the $\b{g}$- and $\b{c}$-vectors.
As defined in~\cite[Def.~3.1]{FominZelevinsky-ClusterAlgebrasIV}, the cluster algebra with \defn{principal coefficients} at~$\B_\circ$ is the cluster algebra~$\clusterAlgebra[\B_\circ][\coefficients_\circ]$ in~$\Q(x_1, \dots, x_n, p_1, \dots, p_n)$, where~$\coefficients_\circ = \{p_x\}_{x \in \cluster_\circ}$ are precisely the generators~$p_1, \dots, p_n$ of~$\Trop{n}$ relabeled by~$\cluster_\circ$.
For simplicity, we drop the mention of the coefficients for principal cluster algebras in the notations for seeds~$\seed = (\B, \cluster)$, for cluster algebras~$\principalClusterAlgebra$ and for variables~$\principalVariables$.
Principal coefficients cluster algebras are $\Z^n$-graded (in the fundamental weight basis~$\{\simpleWeight_x\}_{x \in \cluster_\circ}$ of~$V$) for the degree function~$\deg(\B_\circ,\cdot)$ on $\principalClusterAlgebra$ obtained by setting
\(
\deg(\B_\circ,  x) \eqdef \fundamentalWeight_x
\)
and
\(
\deg(\B_\circ, p_x) \eqdef \sum_{y \in \cluster_\circ} -b_{yx} \fundamentalWeight_y
\)
for any $x\in\cluster_\circ$.

>>>>

\medskip
\enlargethispage{.5cm}
Further families of generalizations of the associahedron certainly deserve to be studied with our type cone approach.
We are investigating in particular graph associahedra~\cite{CarrDevadoss, Postnikov, FeichtnerSturmfels, Zelevinsky}, brick polytopes~\cite{PilaudSantos-brickPolytope, PilaudStump-brickPolytope}\footnote{A preliminary version of this paper contained the full facet description of the type cones of graph associahedra. As they are not simplicial, they lie slightly beyond the scope of this paper, and will be studied elsewhere.} and quotientopes~\cite{Reading-latticeCongruences, Reading-CambrianLattices, PilaudSantos-quotientopes}.
In fact, as already mentioned, it seems sometimes easier to find all polytopal realizations of a fan (by describing its type cone in terms of certain specific pairs of adjacent maximal cones) than to identify one specific explicit realization.
This is clearly the case when the type cone is simplicial as illustrated by the results of this paper: we naturally obtained all polytopal realizations of the $\b{g}$-vector fans of cluster algebras from cyclic seeds and of gentle algebras whose polytopality was only established very recently~\cite{HohlwegPilaudStella, PaluPilaudPlamondon-nonkissing}.
We believe that this approach gives reasonable hope that the polytopality of further fans could in the future be established using our type cone approach.
Tempting candidates are quotientopes for hyperplane arrangements beyond type~$A$, which are particularly interesting since the coarsenings of a fan essentially correspond to the faces of its type cone (see \cref{subsec:facesTypeCone}).

@@

\para{Principal coefficients and $\b{g}$- and $\b{c}$-vectors}
%
We now consider principal coefficients to define the $\b{g}$- and $\b{c}$-vectors.
As defined in~\cite[Def.~3.1]{FominZelevinsky-ClusterAlgebrasIV}, the cluster algebra with \defn{principal coefficients} at~$\B_\circ$ is the cluster algebra~$\clusterAlgebra[\B_\circ][\coefficients_\circ]$ in~$\Q(x_1, \dots, x_n, p_1, \dots, p_n)$, where~$\coefficients_\circ = \{p_x\}_{x \in \cluster_\circ}$ are precisely the generators~$p_1, \dots, p_n$ of~$\Trop{n}$ relabeled by~$\cluster_\circ$.
For simplicity, we drop the mention of the coefficients for principal cluster algebras in the notations for seeds~$\seed = (\B, \cluster)$, for cluster algebras~$\principalClusterAlgebra$ and for variables~$\principalVariables$.
Principal coefficients cluster algebras are $\Z^n$-graded (in the fundamental weight basis~$\{\fundamentalWeight_x\}_{x \in \cluster_\circ}$ of~$V$) for the degree function~$\deg(\B_\circ,\cdot)$ on $\principalClusterAlgebra$ obtained by setting
\(
\deg(\B_\circ,  x) \eqdef \fundamentalWeight_x
\)
and
\(
\deg(\B_\circ, p_x) \eqdef \sum_{y \in \cluster_\circ} -b_{yx} \fundamentalWeight_y
\)
for any $x\in\cluster_\circ$.
